---
title: "Introduction to qldyjcost"
author: "Trent Henderson"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{Introduction to qldyjcost}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 7,
  fig.width = 7,
  warning = FALSE,
  fig.align = "center"
)
```

```{r setup, message = FALSE, warning = FALSE}
library(qldyjcost)
```

## Usage

The core of `qldyjcost` is a data frame of historical costs from peer-reviewed sources called `yjcosts`. It can be accessed easily via:

```{r, message = FALSE, warning = FALSE}
head(yjcosts)
```

`yjcosts` contains historical offending costs broken down by offence seriousness for court costs, police costs, wider social costs, and custody costs.

`qldyjcost` also contains functionality for adjusting the historical offending costs to values for a given financial year using inflation. The `get_cpi` function webscrapes inflation rates from the [Australian Bureau of Statistics](https://www.abs.gov.au/statistics/economy/price-indexes-and-inflation/consumer-price-index-australia/) and automatically cleans it up ready for use:

```{r, message = FALSE, warning = FALSE}
cpi <- get_cpi()
head(cpi)
```

Instead of grabbing CPI for the country, we can instead get it for any capital city:

```{r, message = FALSE, warning = FALSE}
cpi_bris <- get_cpi(location = "Brisbane")
head(cpi_bris)
```

While this function is useful in isolation, it forms a key component of the workhorse function of `yjcosts` which is `adjust_costs`. `adjust_costs` applies inflation adjustments to the offending cost data to any given financial year. Specifically, it does this by averaging CPI values for each FY associated with the baseline costs in `yjcosts` as well as the user-specified FY to adjust to. The interface is incredible simple. Here is how to adjust costs to FY 2023-24 dollars:

```{r, message = FALSE, warning = FALSE}
adj <- adjust_costs(fy = "2023-24", location = "Australia")
head(adj)
```
